\ \\ \LARGE\bf Speculative Evaluation in Particle Swarm Optimization \thanks{Matthew Gardner, Andrew McNabb, and Kevin Seppi are with the Department of Computer Science, Brigham Young University, 3361 TMCB, Provo, UT 84602 (phone: 801-422-8717; email: \{mjg82,a,k\}@cs.byu.edu).}

### 2010-06-09 honors_thesis/journal_paper/paper.tex
Making some modifications from feedback from Chris

Changing the way neighbors are defined, usually called the ``topology,'' has a
significant effect on the performance of the algorithm.  In the Ring topology,
each particle has one neighbor to either side of it; in the Complete topology,
every particle is a neighbor to every other
particle~\cite{bratton-2007-defining-a-standard-for-pso}.  In all topologies a
particle is also a neighbor to itself in that its own position and value are
considered when updating the particle's neighborhood best, $\nbest$.  Thus with
$p$ particles, using the Ring topology each particle with index $i$ has three
neighbors: $i-1$, $i$ (itself), and $i+1$.  With the Complete topology, each
particle has $p$ neighbors.

@@

The PSO algorithm can be trivially parallelized by distributing the computation
needed for each particle on across processors.  But as we have seen in
\fig{evals-sphere}, for some functions, and for large numbers of processors,
just adding particles reaches a point of diminishing returns.  That is, adding
processors does not help the algorithm reach any given level of fitness
appreciably faster.  Instead of adding particles we employ a speculative
approach that allows us to perform iterations two at a time.

@@

In some distributed frameworks, like the MapReduce framework that we used in
our experiments, rounds of message passing can be expensive.  The method just
described uses three rounds of message passing for every two iterations
(corresponding to steps 2, 5 and 10 in \alg{centralized}).  It is possible to
perform speculative evaluation in PSO with only one round of communication per
two iterations.  However, the methodology is tedious and is not the focus of
this paper, so we defer its description to the Appendix.

@@

To make better use of the extra speculative evaluations, instead of choosing
the speculative child that matches the branch that the original PSO would have
taken, we take the child that has the best value.  The methodology is exactly
the same as above except for the process of choosing which speculative child to
accept.  The only change needed in \alg{centralized} is in step 7, where
$\noeval{\s}_{t+1}$ with the best value is chosen from $\noeval{\sset}_{t+1}$
instead of with the matching branch.

This can be thought of as drawing a number of samples from the next iteration
and accepting the best one.  Speculative particles that move in good directions
are kept.  It seems this modification to PSO would produce more exploitation,
decreasing the amount of exploration.  With functions that are less deceptive
we would expect this method to work well, while performance on more deceptive
functions would probably be hurt.  We call this technique Pick Best.

>>>>

Changing the way neighbors are defined, usually called the ``topology,'' has a
significant effect on the performance of the algorithm.  In the Ring topology,
each particle has one neighbor to either side of it; in the Complete
topology\footnote{The Complete topology has often been unfortunately named Star
in the literature, which in graph theory refers to a completely different
topology.  Other names have also been used, including ``global topology'' or
gbest.  We use the graph theory term ``Complete'' in this paper.}, every
particle is a neighbor to every other
particle~\cite{bratton-2007-defining-a-standard-for-pso}.  In all topologies a
particle is also a neighbor to itself in that its own position and value are
considered when updating the particle's neighborhood best, $\nbest$.  Thus with
$p$ particles, using the Ring topology each particle with index $i$ has three
neighbors: $i-1$, $i$ (itself), and $i+1$.  With the Complete topology, each
particle has $p$ neighbors.

@@

PSO can be trivially parallelized by assigning each particle's computation to
an individual processor.  But as we have seen in \fig{evals-sphere}, for some
functions, and for large numbers of processors, just adding particles reaches a
point of diminishing returns.  That is, adding processors does not help the
algorithm reach any given level of fitness appreciably faster.  Instead of
adding particles we employ a speculative approach that allows us to perform
iterations two at a time.

@@

In distributed frameworks, sychronizing all of the machines for a round of
message passing can be expensive.  The method just described uses three rounds
of message passing for every two iterations (corresponding to steps 2, 5 and 10
in \alg{centralized}).  It is possible to perform speculative evaluation in PSO
with only one round of communication per two iterations.  However, the
methodology is tedious and is not the focus of this paper, so we defer its
description to the Appendix.

@@

To make better use of the extra speculative evaluations, instead of choosing
the speculative child that matches the branch that the original PSO would have
taken, we take the child that has the best value.  The methodology is exactly
the same as above except for the process of choosing which speculative child to
accept.  The only change needed in \alg{centralized} is in step 7, where
$\noeval{\s}_{t+1}$ with the best value is chosen from $\noeval{\sset}_{t+1}$
instead of with the matching branch.  We call this technique Pick Best.

This can be thought of as drawing a number of samples from the next iteration
and accepting the best one.  Speculative particles that move in good directions
are kept.  Intuition says that this technique favors exploitation over
exploration, but as we will show in \sec{results}, that is not always the case.
